From Big Data to Big Displays\\ High-Performance Visualization at Blue Brain

### 2017-06-20 paper.tex
Layout and wording changes.

RTNeuron\cite{Hernando13} is the oldest of our interactive visualization tools.
Originally conceived as a standalone application for visualizing simulation
results, it has evolved into a rasterization-based rendering engine library
implemented in C++ with a Python wrapping. The power of RTNeuron lies in a
domain specific API designed for the visualization of detailed neuronal circuit
models, this API allows for building custom applications tailored for specific
use cases. It provides features to visualize static circuit data and some
simulation results. Static data visualization includes different visual
representations for neurons, synapses, selective pruning of neuronal trees,
clipping planes and others. The simulation results that can be displayed are the
spiking activity of the neurons and scalar variables from the cable models
mapped to the neuron surface. The color maps for displaying these data are
highly configurable, allowing to apply different color maps to different cell
subsets.

Neurons can be displayed with different levels of detail ranging from simple
spheres, cylinder-like geometric models and detailed polygonal
meshes. Advanced rendering techniques included are several types of parallel
rendering algorithms (such as sort-first and sort-last) and several algorithms
for transparency rendering that enable efficient and correct rendering of
scenes with great geometrical complexity and high depth
complexity. Transparency is particularly suitable for masking and highlighting
features of interest on the circuit. Ongoing work is a more scalable parallel
rendering algorithm specially suited for transparency. Tiled rendering is also
possible, which allows very high resolution renderings suitable for printing
at sizes larger than A0.

The core engine is implemented in C++ using Equalizer and OpenSceneGraph and
leverages part of our messaging framework to allow coupling with other
applications. The 3D view can be embedded inside Qt applications, in particular
in Python with PyQt and QML for overlaying GUI elements. Several use case
specific applications have been built this way.

@@

\begin{figure}[h!t]
    \includegraphics[width=.45\columnwidth]{images/livre}\hfill
    \includegraphics[width=.45\columnwidth]{images/livreLOD}
  \caption{\label{fLivre}Livre rendering a MicroCT dataset (left) and
    the selected LODs (right)}
\end{figure}

@@

\begin{wrapfigure}{r}{.382\textwidth}\center\vspace{-6ex}
  \includegraphics[width=.382\textwidth]{images/ZeroMS}
  {\caption{\label{fZero}New messaging use cases}}\vspace{-2ex}
\end{wrapfigure}

>>>>

RTNeuron\cite{HBBES:13} is the oldest of our interactive visualization tools.
Originally conceived as a standalone application for visualizing simulation
results, it has evolved into a rasterization-based rendering engine library
implemented in C++ with a Python wrapping. The power of RTNeuron lies in a
domain specific API designed for the visualization of detailed neuronal circuit
models, this API allows for building custom applications tailored for specific
use cases. It provides features to visualize static circuit data and some
simulation results. Static data visualization includes different visual
representations for neurons, synapses, selective pruning of neuronal trees,
clipping planes and others. The simulation results that can be displayed are the
spiking activity of the neurons and scalar variables from the cable models
mapped to the neuron surface. The color maps for displaying these data are
highly configurable, allowing to apply different color maps to different cell
subsets.

Neurons can be displayed with different levels of detail ranging from simple
spheres, cylinder-like geometric models and detailed polygonal
meshes. Advanced rendering techniques included are several types of parallel
rendering algorithms (such as sort-first and sort-last) and several algorithms
for transparency rendering that enable efficient and correct rendering of
scenes with great geometrical complexity and high depth
complexity. Transparency is particularly suitable for masking and highlighting
features of interest on the circuit. Ongoing work is a more scalable parallel
rendering algorithm specially suited for transparency. Tiled rendering is also
possible, which allows very high resolution renderings suitable for printing
at sizes larger than A0 (The original of \fig{fRTNeuron} left is a 36640$\times$26000 pixel image).

The core engine is implemented in C++ using Equalizer and OpenSceneGraph and
leverages part of our messaging framework to allow coupling with other
applications. The 3D view can be embedded inside Qt applications
(\fig{fRTNeuron}, right), in particular in Python with PyQt and QML for
overlaying GUI elements. Several use case specific applications have been built
this way.

\begin{figure}[h!t]\center
  \includegraphics[width=.45\columnwidth]{images/rtneuron1.png}\hfil
  \includegraphics[width=.45\columnwidth]{images/rtneuron2.png}
  \caption{\label{fRTNeuron}RTNeuron examples: membrame potential of a fraction of the neurons in a simulation (left) and a circuit visualization application showing a portion of a hippocampus model (right).}
\end{figure}

@@

\begin{figure}[t]
    \includegraphics[width=.45\columnwidth]{images/livre}\hfill
    \includegraphics[width=.45\columnwidth]{images/livreLOD}
  \caption{\label{fLivre}Livre rendering a MicroCT dataset (left) and
    the selected LODs (right)}
\end{figure}

@@

\begin{wrapfigure}{r}{.382\textwidth}\center\vspace{-6ex}
  \includegraphics[width=.382\textwidth]{images/ZeroMS}
  {\caption{\label{fZero}New messaging use cases}}\vspace{-6em}
\end{wrapfigure}
